### 2018-11-13 papers/journal/encoding.tex
Updated introduction section: Biological part is still in early writing stage. Incomplete update
Added encodings for recahabilty for BMC implementation

In this section we will present the SAT, SMT and QBF based encoding of both the search and the synthesis problem. We will start with the SMT based encoding and later present the contrast with the BMC encoding.

@@

The BMC encoding differs with presented SMT encoding slightly in the way variables are represented and how we encode reachability.  
%
Variables are not Boolean indexed with multiple dimensions but multidimensional arrays. 
%
For clarity, we will only present the stability encoding.

We initialize the system with a fixed number of node size (N) and distinct molecular types (M). 
%
We use a two-dimensional array variable \textit{Graph}[N][N] to represent the transport graph. The graph variable is set to x, Graph[i][j] = x, if there are exactly x edges between node i and j. 
%\begin{align*}
%graph [i] [j] = x.
%\end{align*}

%
The nodes are represented as a bitvectors of size M with each bit corresponding to the presence or the absence of the molecule. 
%
The index of the bitvector implicitly determines the corresponding molecule we are referring to.
%

The molecule set is divided into Q and R-SNAREs. 
%
For clarity, we consider both in the same amount and represent the first half of the bitvector as  Q-SNAREs and rest as R-SNAREs.
%
A node in the system with M = 4 molecules is interpreted as a tag [M1, M2, M3, M4]. 
%
Where M1, M2 are Q-SNAREs and M3, M4 are R-SNAREs.
%
Also, each $M_i$ can be either 0 or 1, representing the presence or absence of the molecule i.
%
Similarly, an edge in the graph is also tagged as a bitvector of length M.   
%
We leave the graph completely arbitrary and add the constraint to enforce that the graph is k-connected.

In order to specify the stability condition, we use a combination of enumeration and nondeterminism.
%
Nondeterminism can be considered as an oracle that guesses a correct $m$-path for a specific m if it exists. 
%
We then use this path as a witness to ensures that stability is maintained for each outgoing molecule m on every edge E.

@@

Now we can encode stability using the NOracle and say if there is an $m$-edge between $i$ and $j$, there is a $m$-path from $j$ to $i$ of length max w.
\begin{align}
 \bigwedge\limits_{i,j,m} (\bigvee_{q} e[i][j][q][m]) \limplies NOracle(m, j, i, w)
  \tag{S2}\label{eq:reachbmc2}
\end{align}

@@

\subsection{SMT problems for different variants}
%\ankit{Bring model variant table here}

>>>>

\externaldocument{sat-table}
\externaldocument{smt-table}

In this section we will present the BMC, SMT and QBF based encoding of both the search and the synthesis problem. We will start with the SMT based encoding and later present the contrast with the BMC encoding.

@@

The BMC encoding differs with the presented SMT encoding slightly in the way variables are represented and how we encode reachability.  
%
Variables are not Boolean indexed with multiple dimensions but arrays of bitvectors of size $\mu$. 
%
%We initialize the system with a fixed number of node size (N) and distinct molecular types (M). 
%%
%We use a two-dimensional array variable \textit{Graph}[N][N] to represent the transport graph. The graph variable is set to x, Graph[i][j] = x, if there are exactly x edges between node i and j. 
%\begin{align*}
%graph [i] [j] = x.
%\end{align*}
%
For example the edges are represented as a bitvectors of size M with each bit corresponding to the presence or the absence of the molecule. 
%
The index of the bitvector implicitly determines the corresponding molecule we are referring to.

\begin{enumerate}
	\item Bit vector $bv\_e[i][j][q]$ represent the $q$th edge between $i$ and $j$.
\end{enumerate}

%

The molecule set is divided into Q and R-SNAREs. 
%
For clarity, we consider both in the same amount and represent the first half of the bitvector as  Q-SNAREs and rest as R-SNAREs.
%
An edge in the system with M = 4 molecules is interpreted as a tag [M1, M2, M3, M4]. 
%
Where M1, M2 are Q-SNAREs and M3, M4 are R-SNAREs.
%
Also, each $M_i$ can be either 0 or 1, representing the presence or absence of the molecule i.
%

\textbf{Example 1} The variable bv\_e[1][2][1] = 1001 represent first edge between node 1 and node 2 with molecule M1 and M4 present.

Similarly, a node in the graph is also tagged as a bitvector of length M.   
%
We leave the graph completely arbitrary and add the constraint to enforce that the graph is k-connected. 
%
For clarity, we will only present the stability encoding.

In order to specify the stability condition, we use a combination of enumeration and nondeterminism.
%
Nondeterminism can be considered as an oracle that guesses a correct $m$-path for a specific m if it exists. 
%
We then use this path as a witness to ensures that stability is maintained for each outgoing molecule m on every edge E.
%

@@

Now we can encode stability using the NOracle and say if there is an $m$-edge between $i$ and $j$, there is a $m$-path from $j$ to $i$ of length max w.
\begin{align}
 \bigwedge\limits_{i,j,m} (\bigvee_{q} (bv\_e[i][j][q] >> m) > 0) \limplies NOracle(m, j, i, w)
  \tag{S2}\label{eq:reachbmc2}
\end{align}

@@

\subsection{BMC and SMT problems for different variants}
%\ankit{Bring model variant table here}

### 2018-11-13 papers/journal/model.tex
Added encodings for recahabilty for BMC implementation

\subsection{BMC and SMT}
$\pairs$ defines which molecules can fuse with which molecules.
%
Let $\pairs(M')$ denote $\{m|(m,m') \in P \text{ and } m' \in M'\}$.
%
$\nodef$ and $\edgef$ are used to define activity of molecules on
nodes and edges respectively.
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
We call $G$ {\em well-structured} if molecules $M$ is divided into
two equal-sized partitions $M_1$ and $M_2$ such that
$P \subseteq M_1 \times M_2$, and
for each $(n,M',n') \in \edges$, $n \neq n'$, 
$M' \subseteq \nlabel(n) \intersection \nlabel(n')$.
%
In other words, a well-structured VTS has no self loops, and 
each edge carry only those molecules that are present in its source
and destination nodes.

@@

\subsection{QBF and synthesis}
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
We say $G$ is {\em fuse compatible} if molecules $M$ is divided into
two partitions $Q$ and $R$ such that
for each $P \in \pairs, |P \intersect Q| = 3 \land |P \intersect R| = 1 $.
%
In other words,
molecules are of two types $Q$ and $R$,
%
pairing relations have sets of four molecules such that three
are of one type (??) and one is of another type (??).
%
% This distinction of molecules is due to section~\ref{mol-types}.
%
We call $G$ {\em well-structured} if it is pair compatible, and
for each $(n,M',n') \in \edges$, $n \neq n'$ and
$M' \subseteq \nlabel(n) \intersection \nlabel(n')$.
%
In addition to fuse compatibility, the well-structured property requires that
here are no self loops, and 
each edge carry only those molecules that are present in its source
and destination nodes.
%
An edge $(n,M',\_) \in \edges$ {\em fuses} with a node $n'$
if there are non-empty set of molecules $M'' \subseteq M'$ and $M''' \subseteq \nlabel(n')$
such that $M''$ are active in the edge, $M'''$ are active in $n'$, and $M'' \union M''' \in \pairs$.
%
We call $G$ {\em well-fused} if each edge $(n,M',n') \in \edges$ fuses
with its destination node $n'$
and can not fuse with any other node.
%
$\pairs$ models biological possible quadruple of molecules that participate in fusion.
%
Each edge must be uniquely supported by such a quadruple of molecules.
%
We also require three of them to be on the node or edge, and the forth one
to be on the other side of the fusion.

% $\pairs(M'')$ are not active in any other node.

@@

\subsection{Uniform}
A {\em path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,\_,n_{i+1}) \in \edges$ for each $ 0 < i < \ell$.
%
For a molecule $m \in M$,
an {\em $m$-path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,M',n_{i+1}) \in \edges$ and $m \in M'$ for
each $ 0 < i < \ell$.
%
A node $n'$ is {\em ($m$-)reachable} from node $n$ in $G$ if there is a ($m$-)path
$n,...,n'$ in $G$.
%
% A node $n'$ is {\em $m$-reachable} from node $n$ in $G$ if there is a
% $m$-path $n,...,n'$ in $G$.
%
We call $G$ {\em stable} if for each $(n,M',n') \in \edges$ and $m \in M'$,
$n$ is $m$-reachable from $n'$.
%
The definition of stability implies that if a molecule leaves a
compartment then there must be a path for it to come back.
%
It ensures that the flow of molecules does not deplete a
compartment.
%
Such a property is often expected in several physical systems
such that electromagnetic fields, fluid dynamics, etc.

@@

Let us discuss rough sizes of the system and expected behaviors.
%
A typical VTS may have about 10 compartments.
%
The VTS may transport a large number of molecules but the molecules that
are relevant of the control of VTS are about 50~\cite{doWeKnowThis}.
%
According to the definition the activity of molecules can be controlled by
the presence of all other molecules.
%
However, in practice the activity of a molecule is controlled by
the presence of ~2-3 other molecules.

>>>>

\textbf{Fusion condition in BMC and SMT:}
$\pairs$ defines which molecules can fuse with which molecules.
%
Let $\pairs(M')$ denote $\{m|(m,m') \in P \text{ and } m' \in M'\}$.
%
$\nodef$ and $\edgef$ are used to define activity of molecules on
nodes and edges respectively.
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
We call $G$ {\em well-structured} if molecules $M$ is divided into
two equal-sized partitions $M_1$ and $M_2$ such that
$P \subseteq M_1 \times M_2$, and
for each $(n,M',n') \in \edges$, $n \neq n'$, 
$M' \subseteq \nlabel(n) \intersection \nlabel(n')$.
%
In other words, a well-structured VTS has no self loops, and 
each edge carry only those molecules that are present in its source
and destination nodes.

@@

\textbf{Fusion condition in QBF and synthesis:}
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
We say $G$ is {\em fuse compatible} if molecules $M$ is divided into
two partitions $Q$ and $R$ such that
for each $P \in \pairs, |P \intersect Q| = 3 \land |P \intersect R| = 1 $.
%
In other words,
molecules are of two types $Q$ and $R$,
%
pairing relations have sets of four molecules such that three
are of one type (??) and one is of another type (??).
%
% This distinction of molecules is due to section~\ref{mol-types}.
%
We call $G$ {\em well-structured} if it is pair compatible, and
for each $(n,M',n') \in \edges$, $n \neq n'$ and
$M' \subseteq \nlabel(n) \intersection \nlabel(n')$.
%
In addition to fuse compatibility, the well-structured property requires that
here are no self loops, and 
each edge carry only those molecules that are present in its source
and destination nodes.
%
An edge $(n,M',\_) \in \edges$ {\em fuses} with a node $n'$
if there are non-empty set of molecules $M'' \subseteq M'$ and $M''' \subseteq \nlabel(n')$
such that $M''$ are active in the edge, $M'''$ are active in $n'$, and $M'' \union M''' \in \pairs$.
%
We call $G$ {\em well-fused} if each edge $(n,M',n') \in \edges$ fuses
with its destination node $n'$
and can not fuse with any other node.
%
$\pairs$ models biological possible quadruple of molecules that participate in fusion.
%
Each edge must be uniquely supported by such a quadruple of molecules.
%
We also require three of them to be on the node or edge, and the forth one
to be on the other side of the fusion.
% $\pairs(M'')$ are not active in any other node.

@@

\vspace{0.2cm}
%\subsection{Uniform}
\noindent
A {\em path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,\_,n_{i+1}) \in \edges$ for each $ 0 < i < \ell$.
%
For a molecule $m \in M$,
an {\em $m$-path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,M',n_{i+1}) \in \edges$ and $m \in M'$ for
each $ 0 < i < \ell$.
%
A node $n'$ is {\em ($m$-)reachable} from node $n$ in $G$ if there is a ($m$-)path
$n,...,n'$ in $G$.
%
% A node $n'$ is {\em $m$-reachable} from node $n$ in $G$ if there is a
% $m$-path $n,...,n'$ in $G$.
%
We call $G$ {\em stable} if for each $(n,M',n') \in \edges$ and $m \in M'$,
$n$ is $m$-reachable from $n'$.
%
The definition of stability implies that if a molecule leaves a
compartment then there must be a path for it to come back.
%
It ensures that the flow of molecules does not deplete a
compartment.
%
Such a property is often expected in several physical systems
such that electromagnetic fields, fluid dynamics, etc.

@@

Let us discuss rough sizes of the system and expected behaviors.
%
A typical VTS may have about 10 compartments.
%
%Across eukaryotes, there are 20 broad SNARE varieties, though individual species can
%contain higher numbers (humans have 41)~\cite{kloepper2007elaborate}..

The VTS may transport a large number of molecules but the molecules that
are \todo{SA: commented line above} relevant of the control of VTS are about 50~\cite{kloepper2007elaborate}.
%
According to the definition the activity of molecules can be controlled by
the presence of all other molecules.
%
However, in practice the activity of a molecule is controlled by
the presence of ~2-3 other molecules.

### 2018-11-13 papers/journal/problem.tex
Added encodings for recahabilty for BMC implementation

\subsection{SMT}
\paragraph{Search problem}
Based upon earlier discussion, we need to answer the following search
question among VTSs.
For a given $k$, size $\nu$, and molecule number $\mu$,
we are searching for well-structured, stable, and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols| = \mu$, and
$G$ is not $k$-connected.

\subsection{BMC}
%The notion of stability and connectivity varies slightly for BMC encoding. 
In this subsection, we will present the concept of stability and connectivity for BMC encoding.

%
We call a graph $G$ {\em stable} if for each $(n,M',n') \in \edges$ and $m \in M'$,
there exists a $m$-path in G with $n_1 = n'$ and $n_\ell = n$.
%

%
We call a graph $G$ {\em connected} if for each $n,n' \in \nodes$,
$n'$ there is a path from $n$ to $n'$ in G.
%
We call a graph $G$ $k$-connected if for each $\edges' \subseteq \edges$ and $|\edges'| < k$,
VTS $(\nodes,\mols,\edges-E',\nlabel,\pairs,\edgef,\nodef)$ is connected.
%

We use $m$-${path}_i^j$ to denote an $m$-path between source node i and target node is j.

For the experimentation, we will restrict ourselves to the elementary cycles (cycles of length one). 
%
This restriction puts an upper limit of N-1 on the length of each searched path. 
%
Creating a formula to search a path of length N-1 for each molecule m on every edge E will be extremely large and harder for the solvers to handle.  
%
Even if there exist shorter paths we have to create the formula of maximum size N-1. 
% 
To resolve these problems we will use a technique called bounded model checking, where the depth of the path can be specified and consequently increased or decreased.  
%
This specified depth is called unwinding in the technical terms.
%and can be specified for each path separately.

\paragraph{Search problem}
Based upon earlier discussion, we need to answer the following search
question among VTSs.
For a given $k$, size $\nu$, molecule number $\mu$, and unwinding depth w,
we are searching for well-structured, stable, and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols| = \mu$, and $G$ is not $k$-connected with w depth of unwinding.

\subsection{QBF and synthesis}

>>>>

\subsection{Existential condition}

Based upon earlier discussion, we need to answer the following search
question among VTSs.
%
Here, we will present both BMC and SMT search problem.

\textbf{SMT Search problem:}
For a given $k$, size $\nu$, and molecule number $\mu$,
we are searching for well-structured, stable, and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols| = \mu$, and
$G$ is not $k$-connected.

\noindent
The notion of stability and connectivity varies slightly for BMC encoding. 
%In this subsection, we will present the concept of stability and connectivity for BMC encoding. 
%
We call a graph $G$ {\em stable} if for each $(n,M',n') \in \edges$ and $m \in M'$,
there exists a $m$-path in G with $n_1 = n'$ and $n_\ell = n$.
%
%
We call a graph $G$ {\em connected} if for each $n,n' \in \nodes$,
$n'$ there is a path from $n$ to $n'$ in G.
%
We call a graph $G$ $k$-connected if for each $\edges' \subseteq \edges$ and $|\edges'| < k$,
VTS $(\nodes,\mols,\edges-E',\nlabel,\pairs,\edgef,\nodef)$ is connected.
%	
We use $m$-${path}_i^j$ to denote an $m$-path between source node i and target node is j.

For the experimentation, we will restrict ourselves to the elementary cycles (cycles of length one). 
%
This restriction puts an upper limit of N-1 on the length of each searched path. 
%
Creating a formula to search a path of length N-1 for each molecule m on every edge E will be extremely large and harder for the solvers to handle.  
%
Even if there exist shorter paths we have to create the formula of maximum size N-1. 
% 
To resolve these problems we will use a technique called bounded model checking, where the depth of the path can be specified and consequently increased or decreased.  
%
This specified depth is called unwinding in the technical terms.

\textbf{BMC Search problem:}
For a given $k$, size $\nu$, molecule number $\mu$, and unwinding depth w,
we are searching for well-structured, stable, and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols| = \mu$, and $G$ is not $k$-connected with w depth of unwinding.

\subsection{QBF and synthesis}
Here we will present the QBF and Synthesis search problem.

### 2018-11-13 papers/journal/property.tex
Added encodings for recahabilty for BMC implementation

\subsubsection{Checking existential conditions} 
In our program, the rules VTS needs to follow 
%
%are we represent the stability, fusion, SNARE pairing and regulation rules, and the connectedness property that the system needs to follow, as 
%
are represented as constrained over the transport graphs (G), label on the nodes and edges and activity of those labels (determined by a Boolean function f). 
%
Each of these constrained functions is defined to be TRUE for a G and f: if and only if the corresponding condition holds for the given G and f. 
%
We use these functions to define the assertion that characterizes the existential condition to be checked. 
%
We will show in later sections how we encode these Boolean functions.

>>>>

In our program, the rules VTS needs to follow 
%
%are we represent the stability, fusion, SNARE pairing and regulation rules, and the connectedness property that the system needs to follow, as 
%
are represented as constrained over the transport graphs (G), label on the nodes and edges and activity of those labels (determined by a Boolean function f). 
%
Each of these constrained functions is defined to be TRUE for a G and f: if and only if the corresponding condition holds for the given G and f. 
%
We use these functions to define the assertion that characterizes the corresponding condition to be checked. 
%
We will show in later sections how we encode these Boolean functions.

\subsubsection{Checking existential conditions}
